Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-19 WSSSPE3/report/wg_main_metrics.tex
Section 5.6-5.8: nanoscopic editing.

Metrics for scientific software are important for tenure and promotion,
scientific impact, discovery, reducing duplication, serving as a basis for
potential industrial interest in adopting software, prioritizing develop and
support towards strategic objectives and making a case for new or continued
funding. However, there is no commonly-used standard for collecting or
presenting metrics, nor is it known if there is a common set of metrics for
scientific software. It is imperative that scientific software stakeholders
understand that it is useful to collect metrics.

@@

\item WSSSPE3 Breakout Session: How can we measure the impact of a code on
research, and its value to the
community?\footnote{\url{https://docs.google.com/document/d/1cgUDH3RxrfsLotWhKKOrXUnaYFhrtjcV1TDRkFtwQKI/edit}}

@@

The group discussion began by agreeing on the common purpose of creating a set
of guidance giving examples of specific metrics for the success of scientific
software in use, why they were chosen, what they are useful to measure, and any
challenges and pitfalls; then publish this as a white paper. The group discussed
many questions related to useful metrics for scientific software including
addressing if there is a common set of metrics that can be filtered in some way,
can metrics be fit into a common template, which metrics would be the most
useful for each stakeholder, which metrics are the most helpful and how would we
assess this, how are metrics monitored, and many more. A more complete bulleted
list of these questions can be found in Appendix H. Next, a roadmap for how to
proceed was discussed including creating a set of milestones and tasks. The idea
was put forth for the group to interact with the organizing committee of the
2016 NSF Software Infrastructure for Sustained Innovation (SI2) PI workshop in
order to send a software metrics survey to all SI2 and related awardees as a
targeted and relevant set of stakeholders. The five solicitations for software
elements released under the NSF SI2 program all included metrics as a required
component with submitters requested to include {\it "a list of tangible metrics,
with end user involvement, to be used to measure the success of the software
element developed, ...'}. These metrics are then reported as part of annual
reports to NSF by the projects. Although neither the proposal text describing
the metrics nor the reported metric results are publicly available, there is
reason to believe that the community will be willing to provide this information
through a survey mechanism. This survey would be created by one of the student
group members. Similarly, it was suggested that a software metrics survey be
sent to the UK SFTF and TRDF software projects to ask them what metrics would be
useful to report. The remainder of the discussion focused mainly on the creation
of a white paper on this topic. This resulted in a paper outline and writing
assignments with the goal of publishing in venues including WSSSPE4, IEEE CISE,
or JORS. More information about the group discussion is available in
Appendix~\ref{sec:appendix_metrics}.

@@

The main plans for the group going forward are the creation of a white paper on
the topic of useful metrics for scientific software. The authoring of this white
paper would happen in parallel with the creation of a survey by the group with
the survey results to be incorporated in the white paper. The timeline for
completion of the white paper is approximately one year targeting venues
discussed in the previous section.

>>>>

Metrics for scientific software are important for tenure and promotion,
scientific impact, discovery, reducing duplication, serving as a basis for
potential industrial interest in adopting software, prioritizing development and
support towards strategic objectives, and making a case for new or continued
funding. However, there is no commonly-used standard for collecting or
presenting metrics, nor is it known if there is a common set of metrics for
scientific software. It is imperative that scientific software stakeholders
understand that it is useful to collect metrics.

@@

\item WSSSPE3 Breakout Session: How can we measure the impact of a piece of code on
research, and its value to the
community?\footnote{\url{https://docs.google.com/document/d/1cgUDH3RxrfsLotWhKKOrXUnaYFhrtjcV1TDRkFtwQKI/edit}}

@@

The group discussion began by agreeing on the common purpose of creating a set
of guidance giving examples of specific metrics for the success of scientific
software in use, why they were chosen, what they are useful to measure, and any
challenges and pitfalls; then publish this as a white paper. The group discussed
many questions related to useful metrics for scientific software including
addressing if there is a common set of metrics that can be filtered in some way,
can metrics be fit into a common template, which metrics would be the most
useful for each stakeholder, which metrics are the most helpful and how would we
assess this, how are metrics monitored, and many more. A more complete bulleted
list of these questions can be found in Appendix H. Next, a roadmap for how to
proceed was discussed including creating a set of milestones and tasks. The idea
was put forth for the group to interact with the organizing committee of the
2016 NSF Software Infrastructure for Sustained Innovation (SI2) PI workshop in
order to send a software metrics survey to all SI2 and related awardees as a
targeted and relevant set of stakeholders. The five solicitations for software
elements released under the NSF SI2 program all included metrics as a required
component with submitters requested to include {\it ``a list of tangible metrics,
with end user involvement, to be used to measure the success of the software
element developed, \dots''}. These metrics are then reported as part of annual
reports to NSF by the projects. Although neither the proposal text describing
the metrics nor the reported metric results are publicly available, there is
reason to believe that the community will be willing to provide this information
through a survey mechanism. This survey would be created by one of the student
group members. Similarly, it was suggested that a software metrics survey be
sent to the UK SFTF and TRDF software projects to ask them what metrics would be
useful to report. The remainder of the discussion focused mainly on the creation
of a white paper on this topic. This resulted in a paper outline and writing
assignments with the goal of publishing in venues including WSSSPE4, IEEE CISE,
or JORS. More information about the group discussion is available in
Appendix~\ref{sec:appendix_metrics}. \choinote{Spell out SFTF, TRDF, CISE, JORS?}

@@

The main plan for the group going forward is the creation of a white paper on
the topic of useful metrics for scientific software. The authoring of this white
paper would happen in parallel with the creation of a survey by the group with
the survey results to be incorporated in the white paper. The timeline for
completion of the white paper is approximately one year targeting venues
discussed in the previous section.

### 2015-11-19 WSSSPE3/report/wg_main_software_credit.tex
minor changes/comments in publishing sections moved some credit text to be a general intro to the section
minor change in start of software credit section
Section 5.6-5.8: nanoscopic editing.

\subsubsection{Why it is important}

Science and engineering research increasingly relies on software, but there is currently no standard mechanism for citing software or receiving credit for developing software akin to receiving credit via citations for writing papers.
Standardizing software citations offers one route to establishing such a mechanism.
Software is currently eligible for DOI assignment, but DOI metadata fields are not well tuned for software compared to publications. 
Some software providers apply for DOIs but it is still not widely adopted. 
Also, there is no mechanism to cite software dependencies within software in the same way papers cite supporting prior work.

@@

A number of groups external to WSSSPE (although with some overlapping members) are also focused on aspects of software credit, including the FORCE11 Software Citation Working Group (see plans for coordination below).
In addition, a Software Credit workshop\footnote{London Software Credit workshop: \url{http://www.software.ac.uk/software-credit}} convened in London on October 19, following the conclusion of WSSSPE3.
See Appendix~\ref{sec:appendix_SW_credit} for more detailed discussion of related activities.

>>>>

%\subsubsection{Why it is important}

Science and engineering research increasingly relies on software, but there is
currently no standard mechanism for citing software or receiving credit for
developing software akin to receiving credit via citations for writing papers.
Standardizing software citations offers one route to establishing such a
citation and credit mechanism. Software is currently eligible for DOI
assignment, but DOI metadata fields are not well tuned for software compared to
publications. Some software providers apply for DOIs but it is still not widely
adopted. Also, there is no mechanism to cite software dependencies within
software in the same way papers cite supporting prior work.

@@

A number of groups external to WSSSPE (although with some overlapping members)
also focused on aspects of software credit, including the FORCE11 Software
Citation Working Group (see plans for coordination below). In addition, a
Software Credit workshop\footnote{London Software Credit workshop:
\url{http://www.software.ac.uk/software-credit}} convened in London on October
19, following the conclusion of WSSSPE3. See
Appendix~\ref{sec:appendix_SW_credit} for more detailed discussion of related
activities.

### 2015-11-19 WSSSPE3/report/wg_main_training.tex
Section 5.6-5.8: nanoscopic editing.

\item WSSSPE Workshop on Sustainable Software for Science: Practice and
Experiences \footnote{\url{http://wssspe.researchcomputing.org.uk/}}
\katznote{change to cite}

\item SEHPCCSE International Workshop on Software Engineering for High
Performance Computing in Computational Science and
Engineering\footnote{\url{http://wssspe.researchcomputing.org.uk/}}
\katznote{fix URL and change to cite}

@@

The main plans for the group are to convene discussion to explore bringing
together a regular meeting of those involved in leading relevant training
projects.

>>>>

\item Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE)\footnote{\url{http://wssspe.researchcomputing.org.uk/}}
\katznote{change to cite}

\item  International Workshop on Software Engineering for High
Performance Computing in Computational Science and
Engineering(SEHPCCSE)\footnote{\url{http://wssspe.researchcomputing.org.uk/}}
\katznote{fix URL and change to cite}

@@

The main plan for the group is to convene a discussion to explore bringing
together regular meetings of those involved in leading relevant training
projects.
